---
title: "STAT/MATH 495: Problem Set 06"
author: "Kiryu Kawahata"
date: "2017-10-17"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)
library(gridExtra)
```





# Collaboration

Please indicate who you collaborated with on this assignment: 





# Setup

Define truth, which again we know for the purposes of this assignment, but in
practice we won't:

* the true function f(x) i.e. the signal
* the true epsilon i.e. the noise, which in this case is Normal$(0, sd=\sigma)$.
Hence the standard deviation $\sigma$ determines the amount of noise.

```{r}
f <- function(x) {
  x^2
}
sigma <- 0.3

```

This is the target point we'll be trying to predict: $(0.95, f(0.95)) = (0.95, 0.95^2) = (0.95, 0.9025)$, Thus, the test set is just `x=0.95`

```{r}
x0 <- 0.95
test_set <- data_frame(x=x0)
```

This function generates a random sample of size $n$; think of this as a "get new
data" function. Random in terms of both:

* (New) the predictor x (uniform on [0,1])
* the amount of noise $\epsilon$

```{r}
generate_sample <- function(f, n, sigma) {
  sample <- data_frame(
    x = runif(n = n, min = 0, max = 1),
    f_x = f(x),
    epsilon = rnorm(n = n, mean = 0, sd = sigma),
    y = f_x + epsilon
  )
  # Recall: We don't observe f(x) and epsilon, just (x, y)
  sample <- sample %>% 
    select(x, y)
  
  return(sample)
}
```

Define

* The number $n$ of observations $(x_i, y_i)$ in each sample. In the handout,
$n=100$ to keep plots uncrowded. Here we boost to $n=500$
* Number of samples of size $n$ to consider

```{r}
n <- 500
n_sample <- 10000


```


# Computation

```{r}
baseplot <- ggplot(data=NULL) +
  geom_vline(xintercept = x0, linetype="dashed") +
  coord_cartesian(xlim=c(0, 1), ylim=c(-0.75, 1.5))
options(expressions = 10000)

plot_df2 <- baseplot + labs(title = "10,000 sampled splines with 2 degrees of freedom")
plot_df99 <- baseplot + labs(title = "10,000 sampled splines with 99 degrees of freedom")
for (i in 1:n_sample) {
  Problem_sample <- generate_sample(f, n, sigma)
  
  Spline_df2 <- smooth.spline(x = Problem_sample$x, y = Problem_sample$y, df = 2) %>%
    broom::augment()
   plot_df2 <- plot_df2 + geom_point(data=Spline_df2, aes(x=x, y=y), alpha=0.1) + geom_line(data=Spline_df2, aes(x=x, y=.fitted), col="blue", alpha=0.1, size=1, )
  
   Spline_df99 <- smooth.spline(x = Problem_sample$x, y = Problem_sample$y, df =99) %>%
    broom::augment()
  
  
   plot_df99 <- plot_df99 + geom_point(data=Spline_df99, aes(x=x, y=y), alpha=0.1) + geom_line(data=Spline_df99, aes(x=x, y=.fitted), col="blue", alpha=0.1, size=1)

}

plot_df2 <- plot_df2 +
  geom_point(data=data_frame(x=x0, y=f(x0)), aes(x=x,y=y), col="red", size=3)
plot_df99 <- plot_df99 +
  geom_point(data=data_frame(x=x0, y=f(x0)), aes(x=x,y=y), col="red", size=3)

Plots_problem <- grid.arrange(plot_df2, plot_df99, nrow=2)




```


# Tables

As done in Lec 2.7, for both

* An `lm` regression AKA a `smooth.splines(x, y, df=2)` model fit 
* A `smooth.splines(x, y, df=99)` model fit 

output tables comparing:

|  MSE| bias_squared|   var| irreducible|   sum|
|----:|------------:|-----:|-----------:|-----:|
|     X|           X  |     X |      X |         X |

where `sum = bias_squared + var + irreducible`. You can create cleanly formatted tables like the one above by piping a data frame into `knitr::kable(digits=4)`.

```{r}


Spline_df2 <- smooth.spline(x = Problem_sample$x, y = Problem_sample$y, df =2) %>%
    broom::augment()


df2_table <- Spline_df2 %>% 
  mutate(MSE = sum(Spline_df2$.resid)^2) %>% 
  mutate(bias_squared = (mean(Spline_df2$fitted) - Spline_df2$y^2)) %>%
  sample_n(1) %>% 
  select(MSE, bias_squared) 
  


```


# Analysis

**Questions**:

1. Based on the topics covered in Lec 2.7, name one possible "sanity check" for your results. Name another if you can.
1. In **two** sentences or less, give a rough sketch of what the procedure would
be to get the breakdown of $$\mbox{MSE}\left[\widehat{f}(x)\right]$$ for *all*
$x$ in this example, and not just for $$\mbox{MSE}\left[\widehat{f}(x_0)\right]
= \mbox{MSE}\left[\widehat{f}(0.95)\right]$$.
1. Which of the two models would you choose for predicting the point of interest and why?

**Answers**:

1.
1.
1. I would choose the 
