---
title: "STAT/MATH 495: Problem Set 06"
author: "Kiryu Kawahata"
date: "2017-10-17"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)
library(gridExtra)
```





# Collaboration

Please indicate who you collaborated with on this assignment: 





# Setup

Define truth, which again we know for the purposes of this assignment, but in
practice we won't:

* the true function f(x) i.e. the signal
* the true epsilon i.e. the noise, which in this case is Normal$(0, sd=\sigma)$.
Hence the standard deviation $\sigma$ determines the amount of noise.

```{r}
f <- function(x) {
  x^2
}
sigma <- 0.3

```

This is the target point we'll be trying to predict: $(0.95, f(0.95)) = (0.95, 0.95^2) = (0.95, 0.9025)$, Thus, the test set is just `x=0.95`

```{r}
x0 <- 0.95
test_set <- data_frame(x=x0) %>% 
  mutate(y = "") 
  
```

This function generates a random sample of size $n$; think of this as a "get new
data" function. Random in terms of both:

* (New) the predictor x (uniform on [0,1])
* the amount of noise $\epsilon$

```{r}
generate_sample <- function(f, n, sigma) {
  sample <- data_frame(
    x = runif(n = n, min = 0, max = 1),
    f_x = f(x),
    epsilon = rnorm(n = n, mean = 0, sd = sigma),
    y = f_x + epsilon
  )
  # Recall: We don't observe f(x) and epsilon, just (x, y)
  sample <- sample %>% 
    select(x, y)
  
  return(sample)
}
```

Define

* The number $n$ of observations $(x_i, y_i)$ in each sample. In the handout,
$n=100$ to keep plots uncrowded. Here we boost to $n=500$
* Number of samples of size $n$ to consider

```{r}
n <- 500
n_sample <- 10000


```


# Computation

```{r}

for (i in 1:n_sample) {
  Problem_sample <- generate_sample(f, n, sigma)
  
  
  Spline_df2 <- smooth.spline(x = Problem_sample$x, y = Problem_sample$y, df = 2)
  
  tidy_spline_df2 <- Spline_df2 %>%   
    broom::augment()

   Spline_df99 <- smooth.spline(x = Problem_sample$x, y = Problem_sample$y, df =99)
    tidy_spline_df99 <- Spline_df99 %>% 
     broom::augment()
  
    Different_y_df2 <- predict(Spline_df2, x = test_set$x) %>% 
  as_tibble() 
    
   }
#Infinite recursion error when attempting to produce plots

Predict_df2 <- predict(Spline_df2, x = test_set$x) %>% 
  as_tibble() 

Predict_df99 <- predict(Spline_df99, x = test_set$x) %>% 
  as_tibble()



```


# Tables

As done in Lec 2.7, for both

* An `lm` regression AKA a `smooth.splines(x, y, df=2)` model fit 
* A `smooth.splines(x, y, df=99)` model fit 

output tables comparing:

|  MSE| bias_squared|   var| irreducible|   sum|
|----:|------------:|-----:|-----------:|-----:|
|     X|           X  |     X |      X |         X |

where `sum = bias_squared + var + irreducible`. You can create cleanly formatted tables like the one above by piping a data frame into `knitr::kable(digits=4)`.

```{r}


table_df2 <- test_set %>%
  mutate(bias = Predict_df2$y - f(x)) %>%
  mutate(bias_squared = bias^2) %>% 
  mutate(var = var(tidy_spline_df2$.resid - f(x))) %>%
  mutate(MSE = (sum(Predict_df2$y - f(x)/nrow(Spline_df2)))^2) %>% 
  mutate(irreducible = sigma^2) %>% 
  mutate(sum = bias_squared+ var+ irreducible) %>% 
  select(MSE, bias_squared, var, irreducible, sum) %>% 
  knitr::kable(digits = 4)
  
table_df99 <- test_set %>%
  mutate(bias = Predict_df99$y - f(x)) %>%
  mutate(bias_squared = bias^2) %>% 
  mutate(var = var(tidy_spline_df99$y - f(x))) %>%
  mutate(MSE = (sum(Predict_df99$y - f(x)/nrow(Spline_df99)))^2) %>% 
  mutate(irreducible = sigma^2) %>% 
  mutate(sum = bias_squared+ var+ irreducible) %>% 
  select(MSE, bias_squared, var, irreducible, sum) %>% 
  knitr::kable(digits = 4)
  
table_df2
table_df99
```


# Analysis

**Questions**:

1. Based on the topics covered in Lec 2.7, name one possible "sanity check" for your results. Name another if you can.
1. In **two** sentences or less, give a rough sketch of what the procedure would
be to get the breakdown of $$\mbox{MSE}\left[\widehat{f}(x)\right]$$ for *all*
$x$ in this example, and not just for $$\mbox{MSE}\left[\widehat{f}(x_0)\right]
= \mbox{MSE}\left[\widehat{f}(0.95)\right]$$.
1. Which of the two models would you choose for predicting the point of interest and why?

**Answers**:

1.One check is to see if the sum of the variance and bias squared is equal to the MSE. I'm having trouble accomplishing this as of writing and will try to fix it or get help.

1. Create a dataset with a range of all values for x, then fit the models however many times using all of these points

1. I don't believe my work here is correct so I can't decide. If I were to go with the results here I would go with the model with 99 degrees of freedom. According to this table the bias squared value is lower as expected, which would mean higher precision being associated with the model. Additionally the variance values are roughly the same between both models, however I'm not sure if this is truly the case. 
